### 2019-06-28 treestats_paper.tex
cor refs

\begin{example}[Phenotypic correlations] \label{ex:site_correlations}
    Suppose that for each sample $u$ we have a numeric phenotype, denoted $z(u)$,
    and we want to compute the correlation between this phenotype
    and the genotype at each site in the tree sequence.
    For convenience, suppose $z$ is normalized so that $\sum_u z(u) = 0$ and $\sum_u z(u)^2 = 1$.
    Then, if $g_j$ is a vector of binary genotypes (so $g_j(u) = 1$ if $u$ carries the derived allele),
    then the covariance of $z$ with $g_j$ is just $\sum_u z(u) g_j(u) = \sum_{u \st g_j(u) = 1} z(u)$,
    i.e., the sum of the phenotypes of all samples carrying the derived allele.
    Since the phenotypes sum to zero, this is also equal to
    $- \sum_{u \st g_j(u) = 0} z(u)$.
    Therefore, if we write $p_j = \sum_u g_j(u)$ for the derived allele frequency,
    we can write the squared correlation as the sum across alleles as
    \begin{align*}
        r_j^2 =
        \frac{\left( \frac{1}{n} \sum_{u \st g_j(u) = 0} z(u)\right)^2}{2p(1-p)}
        + \frac{\left( \frac{1}{n} \sum_{u \st g_j(u) = 1} z(u)\right)^2}{2p(1-p)}  .
    \end{align*}
    We can compute this as a site statistic by defining $\iw_{1}(u) = z(u)$, and $\iw_{2}(u) = 1$,
    and $f(x_1, x_2) = x_1^2 / 2 x_2 (n - x_2)$:
    then $\site(\iw, f)_j = r_j^2$ is the squared correlation between $z$ and the allele at site $j$.
\end{example}

@@

%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Acknowledgements}

>>>>

\begin{example}[Phenotypic correlations] \label{ex:site_correlations}
    Suppose that for each sample $u$ we have a numeric phenotype, denoted $z(u)$,
    and we want to compute the correlation between this phenotype
    and the genotype at each site in the tree sequence.
    For convenience, suppose $z$ is normalized so that $\sum_u z(u) = 0$ and $\sum_u z(u)^2 = 1$.
    Then, if $g_j$ is a vector of binary genotypes (so $g_j(u) = 1$ if $u$ carries the derived allele),
    then the covariance of $z$ with $g_j$ is just $\sum_u z(u) g_j(u) = \sum_{u \st g_j(u) = 1} z(u)$,
    i.e., the sum of the phenotypes of all samples carrying the derived allele,
    divided $\sqrt{n - 1}$, where $n$ is the number of samples.
    Since the phenotypes sum to zero, this is also equal to 
    $- \sum_{u \st g_j(u) = 0} z(u)$.
    Therefore, if we write $p_j = \sum_u g_j(u)$ for the derived allele frequency,
    we can write the squared correlation as the sum across alleles as
    \begin{align*}
        r_j^2 =
        \frac{\left( \sum_{u \st g_j(u) = 0} z(u)\right)^2}{2p(1-p)n(n-1)} 
        + \frac{\left( \sum_{u \st g_j(u) = 1} z(u)\right)^2}{2p(1-p)n(n-1)}  .
    \end{align*}
    We can compute this as a site statistic by defining $\iw_{1}(u) = z(u)$, and $\iw_{2}(u) = 1/n$,
    and $f(x_1, x_2) = x_1^2 / (2 x_2 (1 - x_2) n (n-1))$:
    then $\site(\iw, f)_j = r_j^2$ is the squared correlation between $z$ and the allele at site $j$.
\end{example}

\textbf{Note:} computing correlations with branches instead of SNPs is like stuff done in \citet{zollner2005coalescent} and \citet{minichiello2006mapping}, maybe.

@@

%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Acknowledgements}
Thanks to Graham Coop for useful suggestions.
